Parameter sensitivity and identifiability for a biogeochemical model of hypoxia in the northern {G}ulf of {M}exico\tnoteref{mytitlenote}

### 2017-03-09 ident_manu.tex
much shortening

\title{Parameter sensitivity and identifiability for a biogeochemical model of hypoxia in the northern {G}ulf of {M}exico\tnoteref{mytitlenote}}
\tnotetext[mytitlenote]{Version: Tue Mar 7 16:55:36 2017 -0600, \href{https://github.com/fawda123/identifiability/commit/8f38d5a2d17442eb0f3f3d4db61d379b841c8830}{8f38d5a2d17442eb0f3f3d4db61d379b841c8830}}

@@

The sensitivity of \ac{do} to perturbations of all relevant parameters for the 108 equations was estimated using a five minute timestep of FishTank simulations from January 1\textsuperscript{st} to December 31\textsuperscript{st}, 2006. Irrelevant parameters were removed for several reasons; parameters were not relevant for the \ac{zerod} model (i.e., hydrodynamic parameters), were considered physical constants, or had no effect given initial conditions.  Additionally, FishTank includes six phytoplankton and two zooplankton groups to describe complexity in community structure and foodweb dynamics.  However, structural equations for each group are identical such that chosen parameter values primarily control differences between the groups, e.g., large-bodied or small-bodied plankton, slow-growing or fast-growing plankton, etc.  Initial analyses indicated that parameter sensitivity of dissolved oxygen was identical within the six phytoplankton and zooplankton groups.  To remove obvious redundancies in the model, the sensitivity analyses were conducted using only one phytoplankton and one zooplankton group.  The final parameter set that was evaluated included 51 parameters that were further grouped into one of six categories based on applicable biogeochemical components of the model: optics ($n = $ 4 parameters), organic matter (12), phytoplankton (22), temperature (2), and zooplankton (11).  A full description of the model parameters is available as an appendix in \citetalias{LehrterIP}.

@@

\noindent where $\gamma$ ranges from one to infinity for perfectly identifiable (orthogonal) or unidentifiable (perfectly collinear) results for parameters in a sensitivity matrix $S$.  The sensitivity functions were supplied as a matrix $\hat{S}$ with rows $i$ and columns $j$ (\cref{sijeqn}) that described deviations of predicted \ac{do} from the default parameter values.  Thus, $\gamma$ can be estimated for any subset of parameter combinations using the change in model output for perturbations of individual parameters. Sensitivity matrices were first normalized by dividing by the square root of the summed residuals \citep{Omlin01,Soetaert10}.

The collinearity index $\gamma$ provides a measure of the linear dependence between sensitivity functions (i.e., $S_i$ for $j$ parameters) described above for subsets of parameters. Estimates of $\gamma$ greater than 10-15 suggest parameter sets are poorly identifiable \citep{Brun01,Omlin01}, meaning parameter values that maximize precision on a calibration dataset are inestimable by conventional optimization algorithms given similar effects of the selected parameters on the estimated state variable. Greater sensitivity of a state variable to a subset of parameters does not always imply better identifiability if the effects of individual parameters are similar.  An intuitive interpretation of $\gamma$ is provided by \citet{Brun01} such that a change in a state variable caused by a change in one parameter can be offset by the fraction $1 - 1/\gamma$ by the remaining parameters.  That is, $\gamma = 10$ suggests the relative change in \ac{do} for an arbitrary parameter in the selected set can be compensated for by 90\% with changes in the other parameters.

Initial analyses suggested that considerably limited subsets of parameters were identifiable of the 51 evaluated for the FishTank model.  Given this limitation, parameter selection must consider the competing objectives of increased precision with parameter inclusion and reduced identifability as it relates to optimization.  An additional challenge is the excessively high number of combinations of parameter sets, which complicates selection given sensitivity differences and desired ecological categories of each parameter (e.g., practitioners may only be interested in optics parameters).  For example, \cref{fig:combnex} provides a simple graphic of the unique number of combinations that are possible for different subsets of `complete' parameter sets of different sizes (i.e., based on $n$ choose $k$ combinations equal to $n!/\left(k!\left(n-k\right)!\right)$).  The number of unique combinations increases with the total parameters in the set and is also maximized for moderate selections (e.g., selecting half the total).  For example, over 10$^{14}$ combinations are possible by selecting 25 parameters from a set of 50.  Accordingly, parameter selection is complicated by differing sensitivity, identifiability limits for parameter subsets, and the difficulty of choosing from many combinations.

A set of heuristics was developed that address the tradeoff in model complexity and identifiability given the challenges described above \citep[see also][]{Wagener01}.  These rulesets were developed with the assumption that parameters will be selected with preference for those with high sensitivity and identifability based on $\gamma < 15$ as an acceptable threshold for subsets (e.g., 93\% accountability).  Selection heurestics also recognized that parameter categories (i.e., optics, organic matter, phytoplankton, temperature, zooplankton) may have unequal preferences by model users given questions of interest.  In all selection scenarios, parameters were selected by decreasing sensitivity starting with the most sensitive until identifiability did not exceed $\gamma = 15$ where selections were \begin{inparaenum}[1\upshape)]
\item blocked within parameter category,
\item independent of parameter category,
\item or considering all categories equally.
\end{inparaenum} The selection rules produced seven subsets of parameters that could further be used to optimize model calibration.

\subsection{Extension to other state variables}

The above analyses were repeated for additional state variables estimated by FishTank to provide further descriptions of ecological dynamics that are relevant for hypoxia.  In addition to \ac{do}, other state variables that were evaluated were ammonium, \ac{chla}, irradiance, nitrate, \ac{pom}, \ac{dom}, and phosphorus.  Particulate and dissolved organic matter were estimated as the summation of the respective outputs for organic matter from phytoplankon (\textit{OM1\_A}, \textit{OM2\_A}) and fecal pellets from zooplankton (\textit{OM1\_Z}, \textit{OM2\_Z}, see \citetalias{LehrterIP}).

@@

Local sensitivity analyses showed that \ac{do} was sensitive to perturbations in 38 of the 51 (75\% of total) parameters that were evaluated in FishTank (default panel \cref{fig:sensalltile}, \cref{tab:dosens}). Within each parameter category, \ac{do} was sensitive to three parameters for optics (75\% of all optic parameters), eight for organic matter (67\%), 16 for phytoplankton (73\%), one for temperature (50\%), and 10 for zooplankton (91\%). Although \ac{do} had the greatest sensitivity to parameters in the zooplankton category (as percentage of total), the relative effects varied. Among all parameters, sensitivity values ranged from $L1 = $ $8.34\times 10^{-8}$ for \textit{QminP} (phytoplankton) to $0.05$ for \textit{umax} (phytoplankton), whereas average sensitivity among all parameters was $L1 = $ $9.2\times 10^{-3}$. Within categories (excluding temperature with one sensitive parameter), sensitivity ranged from $4.39\times 10^{-5}$ (\textit{astarOMA}) to $7.51\times 10^{-4}$ (\textit{astar490}) for optics, $4.17\times 10^{-4}$ (\textit{KNH4}) to $6.15\times 10^{-3}$ (\textit{KG1}) for organic matter, $8.34\times 10^{-8}$ (\textit{QminP}) to $0.05$ (\textit{umax}) for phytoplankton, and $3.69\times 10^{-5}$ (\textit{ZQp}) to $0.05$ (\textit{ZKa}) for zooplankton (\cref{tab:dosens}).  Average sensitivity values in each category were $L1 = $ $2.81\times 10^{-4}$ for optics, $2.17\times 10^{-3}$ for organic matter, $0.02$ for temperature, $0.01$ for phytoplankton, and $0.01$ for zooplankton.

@@

\subsection{Subset identifiability}

The identifiability analyses suggested that many parameter subsets exceeded the thresholds of $\gamma = 10, 15$, providing further justification for using selection heuristics for parameter optimization.  Results for \ac{do} are provided first to demonstrate general concepts for the identifiability analyses, followed by an extension to the remaining state variables.  Parameter identifiability for \ac{do} decreased (increasing $\gamma$) at different rates with increasing size of parameter subsets depending on the parameter category or the number of top parameters that were selected (\cref{fig:identplo}). By category, identifiability was lowest for all combinations of parameter subsets in the phytoplankton ($60$\% of subsets less than $\gamma = 15$, $43$\% less than $\gamma = 10$) and zooplankton categories ($53.1$\% less than $\gamma = 15$, $40$\% less than $\gamma = 10$), whereas all combinations were identifiable for optics ($100$\% less than $\gamma = 15, 10$) and a majority identifiable for organic matter ($91.9$\% less than $\gamma = 15$, $76.5$\% less than $\gamma = 10$). Identifiability for parameters in the temperature category was not evaluated because \ac{do} was sensitive to only one parameter (i.e., $\gamma = 1$). Parameter combinations for choosing from the top, top two, top three, and top four parameters in each category together had decreasing identifability with the increasing size of the selection pool (e.g., top one versus top four parameters, \cref{fig:identplo}).  The percentage of parameter subsets that were below the acceptable thresholds for identifability was $100$\% less than $\gamma = 15, 10$ for the top parameter in each category, $90.6$\% and $80.7$\% for the top two, $80.7$\% and $70.9$\% for the top three, and $55.8$\% and $45.7$\% for the top four.  Results for the remaining state variables had similar patterns in identifiability with increasing size of parameter subsets and selection categories, although differences in identifiabiltiy between state variables was observed (\cref{fig:identploall}).  Most notably, nitrate was consistently the least identifiable variable (highest overall $\gamma$), whereas \ac{do} was most identifiable.

@@

An evaluation of the effects of individual parameters on $\gamma$ suggested that some parameters had disproportionate effects on identifiability.  Based on $\gamma = 15$, \cref{fig:identplo} suggests that most parameter sets for organic matter were identifiable, regardless of how many parameters were selected (i.e., two through eight).  However, some subsets were not identifiable such that identification of one or more redundant parameters that are inflating $\gamma$ values could provide useful information.  \Cref{fig:exclex} shows an alternative view of identifiability of \ac{do} with exclusion and inclusion of individual parameters in different sets for the organic matter category.  As before, collinearity increases with more parameters in a subset, although the increase varies depending on which parameter was included or excluded from the set.  For example, inclusion of \textit{KNO3} in a parameter set almost always inflated $\gamma$.  All parameter subsets that did not include \textit{KNO3} were well below $\gamma = 15$, suggesting that removal of this parameter improves identifiability.  Interestingly, the inclusion of some parameters caused a reduction in $\gamma$, which contradicts the general rule that more parameters caused reduced identifiablity.  For example, parameter sets that included \textit{KGcdom} generally had lower $\gamma$ values relative to those that excluded the parameter.

\subsection{Parameter selection}

The above results demonstrated that state variables differed in the magnitudes of sensitivity for each parameter and the number of identifiable subsets, where the latter varied by $\gamma$ thresholds and parameter selection criteria. Results for each of the three selection heuristics (blocked by parameter category, independent of category, all categories equally) applied to each state variable differed in the number of selected parameters and distribution of parameters within each category (\cref{tab:heurist1,tab:heurist2,tab:heurist3}).  In general, a corresponendence was observed between the number of parameters that were selected given the threhold of $\gamma = 15$ and relative identifiability between the state variables.  As noted above, nitrate was the least identifiable variable (\cref{fig:identploall}), whereas other variables (e.g., \ac{do}, irradiance) were more identifiable.  The constraints on identifiability between variables were demonstrated with the selection heuristics.  For example, heuristics for nitrate typically selected only one or two parameters that met the criteria as compared to more identifiable variables that included several parameters. Overall, the first selection heuristic demonstrated that the number of parameters chosen by parameter category differed independently of the state variables (\cref{tab:heurist1}). The number of selected parameters averaged across state variables in decreasing order was 4.25 parameters from the phytoplankton category, 3.5 from organic matter, 2.75 from optics, and 2.38 from zooplankton. The second and third selection heuristics (\cref{tab:heurist2,tab:heurist3}) were similar, although more parameters were generally selected for the third heuristic given equal importance between categories.

\Cref{fig:heurist_stts} demonstrates parameter selection for all state variables following the second heuristic where parameters were chosen by decreasing sensitivity independent of parameter categories (exact values are shown in \cref{tab:heurist2}).  The y-axis shows the relative identifiability with the addition of parameters from one to many on the x-axis (from left to right).  The second to last parameter for each variable is the last parameter selected within the potential threshold of $\gamma = 15$.  Interestingly, the last parameter shown for most of the state variables caused a relatively large increase in $\gamma$ that was disproportionate to the combined identifiability of the preceding parameter set. For most variables, the phytoplankton edibility vector for zooplankton (\textit{ediblevector(Z1)}) caused a dramatic increase in $\gamma$ with inclusion in the parameter set.  In addition to demonstrating the approach for selection with the second heuristic, \cref{fig:heurist_stts} provides an alternative method of identifying parameters that are disproportionately redundant within a given parameter set (i.e., large contribution to $\gamma$).  This infomation is useful if additional parameter selection rules are developed independently of those proposed herein.

@@

Common goals in the application of biogeochemical models of ecosystem processes are to \begin{inparaenum}[1\upshape)]
\item accurately describe the system by matching predictions with observed data \citep{Reckhow90}, and 
\item provide a means of forecasting ecosystem condition with hypothesized management or environmental scenarios \citep{Clark01}. 
\end{inparaenum}
Although these objectives are the focus of most applications, the structural components of process-based models should secondarily provide inference into which ecosystem processes and functions are driving observed changes.  This latter objective represents a more generic scientific purpose of biogeochemical models that extends beyond the applied benefits of describing and predicting change in a particular system.  Modelers often hope to identify universal principles that govern dynamics across systems and the constraint of model parameters to observations provides a means of supporting or refuting hypotheses \citep{Kirchner06}.  Extension of these principles to test the effects of structural changes and observation uncertainty on model predictions provides further information to support validation of model components. This study provided a simple approach to use the effects of parameter perturbations on model state variables to characterize identifiable parameter subsets that vary by parameter selection criteria.  By doing so, we demonstrated that small parameter subsets relative to all sensitive parameters were within the identifiabilty thresholds described in the literature.  The identifiable parameter subsets varied considerably between state variables and the method for parameter selection.  We further demonstrated that changes in the model structure and variation in the initial conditions had an effect on sensitivity which has direct implications for identifiability. In general, these results provide justification for the use of explicit parameter selection heuristics that practitioners should adopt to facilitate model calibration.

Although the results were specific to individual variables, some generalities about hypoxia dynamics can be inferred from the sensitivity analyses. State variables were most sensitive to parameters in the phytoplankton and zooplankton categories, particularly the maximum growth rates (\textit{umax} for phytoplankton, \textit{Zumax} for zooplankton), mortality coefficient for phytoplankton (\textit{mA}), and the zooplankton half saturation coefficent for grazing (\textit{ZKa}). An increase in the growth rate of primary producers has the potential to increase oxygen concentration through photosynthetic processes, although increased production of organic matter is balanced with respiration and bacterial decomposition that reduce \ac{do} in the water column.  Similarly, increases in zooplankton abundance with increased growth rates causes a reduction in phytoplankton biomass through grazing, which is expected to further deplete pools of organic matter.  Most variables were also sensitive to variation in the half-saturation grazing coefficient which moderates the concentration of nutrients that support half the maximum grazing rate. Although the tradeoff between abundance, grazing, and decomposition is complex, the sensitivity of model state variables to parameters that directly control the abundance of primary producers is in agreement with empirical observations of factors that influence hypoxia dynamics on the \ac{lcs} \citep{Fahnenstiel95,Roelke00,Eldridge10}. The sensitivity of the model output to variation in other parameters that relate to physical and chemical properties of the system was of secondary importance to biological relationships.  That is, state variables were sensitive to changes in light and temperature parameters, although to a lesser extent than phytoplankton and zooplankton parameters.  As such, the differing sensitivities of state variables to parameters in each of the categories was not unexpected given general ecological relationships that are well understood and described by the model.

The overwhelming conclusion from the identifiability analyses is that only limited subsets of parameters were identifiable within the constraints of local sensitivity analyses. Although we have not attempted actual model calibration (see recommendations below), these results support previous studies that have suggested similarly small subsets of parameters can be identified using traditional calibration schemes \citep[e.g.,][]{Wheater86,Ye97,Omlin01}.  In addition to \ac{cgem}, these conclusions have relevance for many biogeochemical models that include numerous parameters and structural equations to characterize processes in the model domain.  A general conclusion is that perhaps a trend towards less complex models could be beneficial given that only a small subset of parameters is identifiable and that ecosystem processes may in fact be sufficiently characterized with few parameters \citep{Ye97}.  Conversely, others have argued that model complexity is not in itself a disadvantage when parsimony is not the sole arbitrator of model structure \citep{Reichert97}. Over-parameterization can be useful if processes have importance that were not evaluated during model identification.  Single objective functions that maximize model precision with identifiable parameters may also provide an incomplete characterization of model worth, which has prompted the development of probability-based models of hypoxia that explicitly include uncertainty in model components \citep[e.g.][]{Obenour15}.  Our results demonstrated that approximately 75\% of the evaluated parameters had an effect on the eight state variables, whereas \ac{cgem} includes a total of 36 variables and multiple plankton groups, not all of which have immediate concern for understanding hypoxia. The redundancies identified with the sensitivity analyses are only problematic if the primary interest is, for example, \ac{do} dynamics.  Moreover, the proposed selection heuristics provide flexibility for choosing different parameters with the assumption that those chosen depend on the research or management question.

Results from the identifiability analyses provided additional insight into the interactions of parameters in large biogeochemical models.  First, identifiability of parameter subsets was not related to the sensitivity of individual variables. As noted above, an identifiable parameter is one that has a unique effect on model predictions that cannot be compensated for or undone by changing other parameters. The magnitude of the effect of a parameter has no bearing on identifiability, which further complicates the selection of parameters for calibration.  Although identifiability is the primary limiting factor in choosing a set, the relative sensitivities are more important for the decision to include or exclude individual parameters.  Our analysis addressed this challenge by presenting multiple selection criteria for identifiable parameter sets that prioritized the most sensitive parameters during the selection process.  Similarly, identifability was not always related to the number of parameters in a set. Although the overwhelming trend was decreasing identifiability with more parameters, the unique effects of including an individual parameter with an existing set often reduced the $\gamma$ estimate. For example, \cref{fig:exclex} shows that including \textit{KGcdom}, \textit{KO2}, or \textit{nitmax} in parameter sets more often reduced $\gamma$ relative to sets that excluded the parameters.  Conversely, \cref{fig:heurist_stts} showed that the inclusion of a specific parameters often caused a disproportionate increase in $\gamma$ (e.g., \textit{ediblevector(Z1)}). These examples demonstrate the complex interactions of parameter changes on variable response, highlighting the need to consider the combined and individual effects of parameters on identifiability. The selection criteria proposed in our above analyses can facilitate parameter selection and also provide diagnostic tools to identify parameters with disproportionate effects on $\gamma$.

@@

An evaluation of parameter uncertainty and identifiability of relevant parameter sets is a preliminary and simplistic approach to improving model predictions.  In general, uncertainty analyses that lead to improved models are ultimately expected to increase our understanding of properties defining ecological relationships.  The extension of simple parameter sensitivity analyses to the generalization of ecosystem properties requires additional analysis and potential model refinement, at the core of which is the balance between generality and precision. We emphasize that the utility of a specific model depends on the question and objective for application to a specific system.  For the above analysis, the FishTank model, as part of the larger \ac{cgem} application, was evaluated in the context of hypoxia effects on ecosystem condition and function.  At the core of the simple model is a set of biogeochemical equations that characterize the system in relation to planktonic groups, nutrient requirements, and water-column metabolism.  Our results have shown that relatively small subsets of parameters are identifiable given the complexity of the model, and as a result, we have provided a general approach to select parameter subsets depending on the ecological context (i.e., selection by parameter category, selection for specific state variables).  Thus, the results described above have relevance for further model refinement with the specific goal of better understanding ecological dynamics that moderate hypoxia on the northern \ac{gom}.  However, the general principles of sensitivity and parameter identifiability have broad applicability beyond this context and we argue that such methods should be more universally applied as an initial approach to quantify numerical constraints of biogeochemical models.

>>>>

\title{Parameter sensitivity and identifiability for a biogeochemical model of hypoxia in the northern {G}ulf of {M}exico\tnoteref{mytitlenote}}
\tnotetext[mytitlenote]{Version: Wed Mar 8 10:38:46 2017 -0600, \href{https://github.com/fawda123/identifiability/commit/f654ac2857b352ca61eae17b12e5450b982e6825}{f654ac2857b352ca61eae17b12e5450b982e6825}}

@@

The sensitivity of state variables to perturbations of all relevant parameters for the 108 equations was estimated using a five minute timestep from January 1\textsuperscript{st} to December 31\textsuperscript{st}, 2006. Irrelevant parameters were removed for several reasons; parameters were not relevant for the \ac{zerod} model (i.e., hydrodynamic parameters), were considered physical constants, or had no effect given initial conditions.  Additionally, FishTank includes six phytoplankton and two zooplankton groups to add complexity in community structure and foodweb dynamics. To remove obvious redundancies, the sensitivity analyses were conducted using only one phytoplankton and one zooplankton group.  The final set that was evaluated included 51 parameters that were further grouped into one of six categories based on applicable biogeochemical components of the model: optics ($n = $ 4 parameters), organic matter (12), phytoplankton (22), temperature (2), and zooplankton (11).  A full description of the model parameters is available as an appendix in \citetalias{LehrterIP}.

@@

\noindent where $\gamma$ ranges from one to infinity for perfectly identifiable (orthogonal) or unidentifiable (perfectly collinear) parameter sets.  The sensitivity functions were supplied as a matrix $\hat{S}$ with rows $i$ and columns $j$ (\cref{sijeqn}) that described deviations of predicted \ac{do} following perturbations of each parameter.  Thus, $\gamma$ can be estimated from results for any subset of parameter combinations. Sensitivity matrices were first normalized by dividing by the square root of the summed residuals \citep{Omlin01,Soetaert10}. Estimates of $\gamma$ greater than 10-15 suggest parameter sets are poorly identifiable \citep{Brun01,Omlin01}, meaning parameter values that maximize precision on a calibration dataset are inestimable by conventional optimization algorithms. An intuitive interpretation of $\gamma$ is provided by \citet{Brun01}, such that a change in a state variable caused by a change in one parameter can be offset by the fraction $1 - 1/\gamma$ by the remaining parameters.  That is, $\gamma = 10$ suggests the relative change in \ac{do} for an arbitrary parameter in the selected set can be compensated for by 90\% with changes in the other parameters.

Parameter selection for model calibration must consider the competing objectives of increased precision with parameter inclusion and reduced identifability as it relates to optimization.  An additional challenge is a large number of combinations of parameter sets, which complicates selection given sensitivity differences and desired ecological categories of each parameter (e.g., practitioners may only be interested in optics parameters).  \cref{fig:combnex} provides a simple graphic of the unique number of combinations that are possible for different subsets of `complete' parameter sets of different sizes (i.e., $n$ choose $k$ combinations, $n!/\left(k!\left(n-k\right)!\right)$).  The number of unique combinations increases with the total parameters in the set and is also maximized for moderate selections (e.g., selecting half the total).  For example, over 10$^{14}$ combinations are possible by selecting 25 parameters from a set of 50.

A set of heuristics was developed that addresses the tradeoff in model complexity and identifiability given the challenges described above \citep[see also][]{Wagener01}.  These rulesets were developed with the assumption that parameters will be selected with preference for those with high sensitivity and identifability based on $\gamma < 15$ as an acceptable threshold for subsets (e.g., 93\% accountability between parameters).  Selection heurestics also recognized that parameter categories (i.e., optics, organic matter, phytoplankton, temperature, zooplankton) may have unequal preferences by model users given questions of interest.  In all selection scenarios, parameters were selected by decreasing sensitivity starting with the most sensitive until identifiability did not exceed $\gamma = 15$ where selections were \begin{inparaenum}[1\upshape)]
\item blocked within parameter category,
\item independent of parameter category,
\item or considering all categories equally.
\end{inparaenum} The selection rules produced seven subsets of parameters that could further be used to optimize model calibration.

@@

Local sensitivity analyses showed that \ac{do} was sensitive to perturbations in 38 of the 51 (75\% of total) parameters that were evaluated in FishTank (\cref{fig:sensalltile}, \cref{tab:dosens}). Within each parameter category, \ac{do} was sensitive to three parameters for optics (75\% of all optic parameters), eight for organic matter (67\%), 16 for phytoplankton (73\%), one for temperature (50\%), and 10 for zooplankton (91\%). Although \ac{do} had the greatest sensitivity to parameters in the zooplankton category (as percentage of total), the relative effects varied. Among all parameters, average sensitivity was $L1 = $ $9.2\times 10^{-3}$ with values ranging from $L1 = $ $8.34\times 10^{-8}$ for \textit{QminP} (phytoplankton) to $0.05$ for \textit{umax} (phytoplankton). Within categories (excluding temperature with one sensitive parameter), sensitivity ranged from $4.39\times 10^{-5}$ (\textit{astarOMA}) to $7.51\times 10^{-4}$ (\textit{astar490}) for optics, $4.17\times 10^{-4}$ (\textit{KNH4}) to $6.15\times 10^{-3}$ (\textit{KG1}) for organic matter, $8.34\times 10^{-8}$ (\textit{QminP}) to $0.05$ (\textit{umax}) for phytoplankton, and $3.69\times 10^{-5}$ (\textit{ZQp}) to $0.05$ (\textit{ZKa}) for zooplankton (\cref{tab:dosens}).  Average sensitivity values in each category were $L1 = $ $2.81\times 10^{-4}$ for optics, $2.17\times 10^{-3}$ for organic matter, $0.02$ for temperature, $0.01$ for phytoplankton, and $0.01$ for zooplankton.

@@

\subsection{Identifiability of parameter subsets and selection rules}

The identifiability analyses suggested that many parameter subsets exceeded the thresholds of $\gamma = 10, 15$.  Parameter identifiability for \ac{do} decreased (increasing $\gamma$) at different rates with increasing size of parameter subsets depending on the parameter category or the number of top parameters that were selected (\cref{fig:identplo}). By category, identifiability was lowest for all combinations of parameter subsets in the phytoplankton ($60$\% of subsets less than $\gamma = 15$, $43$\% less than $\gamma = 10$) and zooplankton categories ($53.1$\% less than $\gamma = 15$, $40$\% less than $\gamma = 10$), whereas all combinations were identifiable for optics ($100$\% less than $\gamma = 15, 10$) and a majority identifiable for organic matter ($91.9$\% less than $\gamma = 15$, $76.5$\% less than $\gamma = 10$). Identifiability for parameters in the temperature category was not evaluated because \ac{do} was sensitive to only one parameter (i.e., $\gamma = 1$). Parameter combinations for choosing from the top, top two, top three, and top four parameters in each category together had decreasing identifability with the increasing size of the selection pool (e.g., top one versus top four parameters, \cref{fig:identplo}).  The percentage of parameter subsets that were below the acceptable thresholds for identifability was $100$\% less than $\gamma = 15, 10$ for the top parameter in each category, $90.6$\% and $80.7$\% for the top two, $80.7$\% and $70.9$\% for the top three, and $55.8$\% and $45.7$\% for the top four.  Results for the remaining state variables had similar patterns in identifiability with increasing size of parameter subsets and selection categories, although differences in identifiabiltiy between state variables was observed (\cref{fig:identploall}).  Most notably, nitrate was consistently the least identifiable variable (highest overall $\gamma$), whereas \ac{do} was most identifiable.

@@

An evaluation of the effects of individual parameters on $\gamma$ suggested that some parameters had disproportionate effects on identifiability.  Based on $\gamma = 15$, \cref{fig:identplo} suggests that most parameter sets for organic matter were identifiable, regardless of how many parameters were selected (i.e., two through eight).  However, some subsets were not identifiable such that identification of one or more redundant parameters that are inflating $\gamma$ values could provide useful information.  \Cref{fig:exclex} shows an alternative view of identifiability of \ac{do} with exclusion and inclusion of individual parameters in different sets for the organic matter category.  As before, collinearity increases with more parameters in a subset, although the increase varies depending on which parameter was included or excluded from the set.  For example, inclusion of \textit{KNO3} in a parameter set almost always inflated $\gamma$.  All parameter subsets that did not include \textit{KNO3} were well below $\gamma = 15$, suggesting that exclusion of this parameter improves identifiability.  Interestingly, the inclusion of some parameters caused a reduction in $\gamma$, which contradicts the general rule that more parameters caused reduced identifiablity.  For example, parameter sets that included \textit{KGcdom} generally had lower $\gamma$ values relative to those that excluded the parameter.

Results for each of the three selection heuristics (blocked by parameter category, independent of category, all categories equally) applied to each state variable differed in the number of selected parameters and distribution of parameters within each category (\cref{tab:heurist1,tab:heurist2,tab:heurist3}).  In general, a corresponendence was observed between the number of parameters that were selected given the threhold of $\gamma = 15$ and relative identifiability between the state variables.  As noted above, nitrate was the least identifiable variable (\cref{fig:identploall}), whereas other variables (e.g., \ac{do}, irradiance) were more identifiable.  The constraints on identifiability between variables were demonstrated with the selection heuristics.  For example, heuristics for nitrate typically selected only one or two parameters that met the criteria as compared to more identifiable variables that included several parameters. Overall, the first selection heuristic demonstrated that the number of parameters chosen by parameter category differed independently of the state variables (\cref{tab:heurist1}). The number of selected parameters averaged across state variables in decreasing order was 4.25 parameters from the phytoplankton category, 3.5 from organic matter, 2.75 from optics, and 2.38 from zooplankton. The second and third selection heuristics (\cref{tab:heurist2,tab:heurist3}) were similar, although more parameters were generally selected for the third heuristic given equal importance between categories.

@@

Common goals in the application of biogeochemical models of ecosystem processes are to \begin{inparaenum}[1\upshape)]
\item accurately describe the system by matching predictions with observed data \citep{Reckhow90}, and 
\item provide a means of forecasting ecosystem condition with hypothesized management or environmental scenarios \citep{Clark01}. 
\end{inparaenum}
Although these objectives are the focus of most applications, the structural components of process-based models should secondarily provide inference into which ecosystem processes and functions are driving observed changes.  This latter objective represents a more generic scientific purpose of biogeochemical models that extends beyond the applied benefits of describing and predicting change in a particular system.  Modelers often hope to identify universal principles that govern dynamics across systems and the constraint of model parameters to observations provides a means of supporting or refuting hypotheses \citep{Kirchner06}.  Extension of these principles to test the effects of structural changes and observation uncertainty on model predictions provides further information to support validation of model components. This study provided a simple approach to use the effects of parameter perturbations on model state variables to characterize identifiable parameter subsets that vary by parameter selection criteria.  By doing so, we demonstrated that small parameter subsets relative to all sensitive parameters were within the identifiability thresholds described in the literature.  The identifiable parameter subsets varied considerably between state variables and the method for parameter selection.  In general, these results provide justification for the use of explicit parameter selection heuristics that practitioners should adopt to facilitate model calibration.

State variables were most sensitive to parameters in the phytoplankton and zooplankton categories, particularly the maximum growth rates (\textit{umax} for phytoplankton, \textit{Zumax} for zooplankton), mortality coefficient for phytoplankton (\textit{mA}), and the zooplankton half saturation coefficent for grazing (\textit{ZKa}). An increase in the growth rate of primary producers has the potential to increase oxygen concentration through photosynthetic processes, although increased production of organic matter is balanced with respiration and bacterial decomposition that reduce \ac{do} in the water column.  Similarly, increases in zooplankton abundance with increased growth rates causes a reduction in phytoplankton biomass through grazing, which is expected to further deplete pools of organic matter.  Most variables were also sensitive to variation in the half-saturation grazing coefficient which moderates the concentration of nutrients that support half the maximum grazing rate. Although the tradeoff between abundance, grazing, and decomposition is complex, the sensitivity of model state variables to parameters that directly control the abundance of primary producers is in agreement with empirical observations of factors that influence hypoxia dynamics on the \ac{lcs} \citep{Fahnenstiel95,Roelke00,Eldridge10}. The sensitivity of the model output to variation in other parameters that relate to physical and chemical properties of the system was of secondary importance to biological relationships.  That is, state variables were sensitive to changes in light and temperature parameters, although to a lesser extent than phytoplankton and zooplankton parameters.  As such, the differing sensitivities of state variables to parameters in each of the categories was not unexpected given general ecological relationships that are well understood and described by the model.

A general conclusion from the identifiability analyses is that only limited subsets of parameters were identifiable within the constraints of local sensitivity analyses. These results support previous studies that have suggested similarly small subsets of parameters can be identified using traditional calibration schemes \citep[e.g.,][]{Wheater86,Ye97,Omlin01}.  In addition to \ac{cgem}, these conclusions have relevance for many biogeochemical models that include numerous parameters and structural equations to characterize processes in the model domain.  A general conclusion is that perhaps a trend towards less complex models could be beneficial given that only a small subset of parameters is identifiable and that ecosystem processes may in fact be sufficiently characterized with few parameters \citep{Ye97}.  Conversely, others have argued that model complexity is not in itself a disadvantage when parsimony is not the sole arbitrator of model structure \citep{Reichert97}. Over-parameterization can be useful if processes have importance that were not evaluated during model identification.  Single objective functions that maximize model precision with identifiable parameters may also provide an incomplete characterization of model worth, which has prompted the development of probability-based models of hypoxia that explicitly include uncertainty in model components \citep[e.g.][]{Obenour15}.  Our results demonstrated that approximately 75\% of the evaluated parameters had an effect on the eight state variables, whereas \ac{cgem} includes a total of 36 variables and multiple plankton groups, not all of which have immediate concern for understanding hypoxia. The redundancies identified with the sensitivity analyses are only problematic if the primary interest is, for example, \ac{do} dynamics.  Moreover, the proposed selection heuristics provide flexibility for choosing different parameters with the assumption that those chosen depend on the research or management question.

Results from the identifiability analyses provided additional insight into the interactions of parameters in large biogeochemical models.  First, identifiability of parameter subsets was not related to the sensitivity of individual variables. As noted above, an identifiable parameter is one that has a unique effect on model predictions that cannot be compensated for or undone by changing other parameters. The magnitude of the effect of a parameter has no bearing on identifiability, which further complicates the selection of parameters for calibration.  Although identifiability is the primary limiting factor in choosing a set, the relative sensitivities are more important for the decision to include or exclude individual parameters.  Our analysis addressed this challenge by presenting multiple selection criteria for identifiable parameter sets that prioritized the most sensitive parameters during the selection process.  Similarly, identifability was not always related to the number of parameters in a set. Although the overwhelming trend was decreasing identifiability with more parameters, the unique effects of including an individual parameter with an existing set often reduced the $\gamma$ estimate. For example, \cref{fig:exclex} shows that including \textit{KGcdom}, \textit{KO2}, or \textit{nitmax} in parameter sets more often reduced $\gamma$ relative to sets that excluded the parameters.  These examples demonstrate the complex interactions of parameter changes on variable response, highlighting the need to consider the combined and individual effects of parameters on identifiability. The selection criteria proposed in our above analyses can facilitate parameter selection and also provide diagnostic tools to identify parameters with disproportionate effects on $\gamma$.

@@

An evaluation of sensitivity and identifiability of relevant parameter sets is a preliminary and simplistic approach to improving model predictions.  In general, uncertainty analyses that lead to improved models are ultimately expected to increase our understanding of properties defining ecological relationships.  The extension of simple parameter sensitivity analyses to the generalization of ecosystem properties requires additional analysis and potential model refinement, at the core of which is the balance between generality and precision. The utility of a specific model depends on the question and objective for application to a specific system.  For the above analysis, the FishTank model, as part of the larger \ac{cgem} application, was evaluated in the context of hypoxia effects on ecosystem condition and function.  At the core of the simple model is a set of biogeochemical equations that characterize the system in relation to planktonic groups, nutrient requirements, and water-column metabolism.  Our results have shown that relatively small subsets of parameters are identifiable given the complexity of the model, and as a result, we have provided a general approach to select parameter subsets depending on the ecological context (i.e., selection by parameter category, selection for specific state variables).  Thus, the results described above have relevance for further model refinement with the specific goal of better understanding ecological dynamics that moderate hypoxia on the northern \ac{gom}.  However, the general principles of sensitivity and parameter identifiability have broad applicability beyond this context and we argue that such methods should be more universally applied as an initial approach to quantify numerical constraints of biogeochemical models.
